HealpixNet: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2018-08-10 scnn.tex
related work in good shape

\section*{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

	\item (MD) distinction between CNN and FCN
    \item (done!) (NP) make figure A.13 remove white space
    \item (first version done!)(NP+MD) paragraph on our chosen architecture in the Experiments sections
    \item (Done) (NP) vertex interpretation 
    \item (MD) Are you happy with the vertex interpretation in section 2.5 Efficient convolutions? Also do you like the end of 2.4?
    \item (NP) Finalise readme on Github code
    \item branch with notebooks with plots (results) NP
    \item NP move the repo to HealpixNet
    \item check if papers which appear as arxiv were not published in a conference or journal
    \item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings. \nati{I think this is too much for this paper.}
    \item TK check part sphere notebook and whole sphere
    \item (done!) (TK) update the README for data zip - file from NP
    \item (almost done) TK read everything
    \item (done!) (TK) uniformize text size in figures
    \item (done!) (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook. \nati{We need to talk about that.}
    \item (done!) (TK) re-make figures to make the labels/legend/titles the same font size as the main text
    \item (done!) TK check the appendix
    \item (done!) TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item (done!) (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (done!) TK read the result section
    \item (done!) (TK) add reference to figure (pooling) to the area of orders
    \item (done!) (TK) add section in conclusions/ paragraph on possible extensions to the code: dense output, regression, GANs, segmentation
	\item Beware to not confuse reader with synonyms, e.g. pixelization and sampling.
    \item Check names: HealpixNet, HealpixCNN, HealpixFCN
    \item Be sure that the convolution on SO(3) is clear: we want to say (a) is the one associated with the SO(3) rotation group (b) can be performed  in the spectral domain by a multiplication (c) is equivariant to rotation (d) is described well in Section 2 of \citep{kondor2018clebsch}.
	\item Have a "Contributions" paragraph?
	\item Notation: write element $i$ of vector $y$ as $y_i$, $y(i)$, or $y[i]$? Be consistent. Same for matrices.
    % \item (TK took exec decision and moved it :D) Should the heat example go to appendix?
    % \item (done! first version done) conclusions
    % \item (done!) Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    % \item (done!) (TK NP) if we have time make a new candidate for network plot (optional)
    % \item (done!) TK read the heat section again
    % \item (done!) (TK) make a new proposition for a plot Figure 6: Two levels of coarsening and pooling
    % \item \todo{Notation GC32 - P256 - FC10 to describe an architecture? Maybe more useful in Experiments where we describe the architecture used for each experiment.}
    % Nati: I did it without this notation. Since we have two architectures, I think it is OK.
\end{itemize}

@@

% \todo{@Tomek: can you explain why we care about parts of the sphere? (that is an advantage of the graph model over the spherical harmonics)}
% \todo{@Tomek: How is the data acquired? Do telescopes and satellites take measurements according to the HEALPix sampling? (it is also an advantage of our method to be agnostic to the sampling, which might be useful to researchers working with missing data or different models)}
% \nati{@Tomek: I think what Michael has in mind is that the graph approach is very versatile and extendable. It can be used with other samplings (even irregular ones) or only part of the sphere. This flexibility might be an advantage to mention. In particular, compared to~\citet{cohen2018spherical} it allows us to do only do the required computation when only a part of the sphere is considered.}

So far these algorithms have mostly been demonstrated on Euclidean domains, such as flat images.
Perhaps the most straightforward approach to the analysis of spherical data is to divide it into small chunks and project those on flat 2D surfaces.
While this approach has been demonstrated to work \citep{fluri2018deep,gupta2018nongaussianinformation,schmelze2017cosmologicalmodel,gillet2018deeplearning}, it may be more natural for some applications and accurate to work directly on the sphere.
The main challenge in designing a CNN on the sphere is to define a convolution operation that is suitable for this irregularly sampled domain.
To be able to train the network efficiently, the convolution has to be performed fast.
Another important feature is for a CNN to be able to work on part-spheres, as for many cosmological observations cover only a part of the sky.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location, and for space-based instruments due to masking of the galactic plane area.

% cohen2017convolutional (ICML workshop) and cohen2018spherical (ICLR best paper) are redundant
A first attempt to generalize CNNs to spherical data was to apply a standard 2D CNN to a grid discretisation of the sphere~\citep{boomsma2017spherical}. While this approach uses the well-developed 2D convolution and hierarchical pooling, it is restricted to those specific grid samplings which are highly irregular on the sphere.
An attempt by~\citet{cohen2018spherical} was to leverage the spherical Fourier transform and to perform the SO(3) convolution in the spectral domain, thanks to the convolution theorem. 
The resulting convolution is rotational equivariant, meaning that a rotation of the input implies the same rotation of the output.
While there exists fast spherical Fourier transforms for some samplings, this approach can still be computationally expensive.
\todo{Add here the paper of Risi?}
Finally, the use of a graph to model the discretised sphere was considered by~\citet{khasanova2017graphomni} as well, towards the goal of designing network that were invariant to rotations.
While their architecture accommodates these invariances on the sphere, they did not take advantage of a hierarchical pixelisation.
Moreover, the statistical layer, which introduces the invariance, needs to be hand-crafted to capture statistics of interest.

To overcome those shortcomings, we introduce an efficient spherical CNN formulation which uses (i) convolutions on graphs and (ii) hierarchical pooling.
The main idea is to represent the sphere $S^2$, a 2-dimensional manifold embedded in $\mathbb{R}^3$, as a graph of connected pixels.
The length (or weight) of the shortest path between two pixels is an approximation of the geodesic distance between them on the manifold.
The use of this approach is enabled by recent developments in graph signal processing, where fundamental tools, such as for example the Fourier transform, were defined \citep{shuman2013emerging} for graphs.
Our approach uses the graph CNN formulation introduced by \citet{defferrard2016convolutional}, which can be adapted to any sampling of the sphere.
Using a $k$-nearest neighbours graph, the convolution operation costs $\mathcal{O}(N_{pix})$ operations, where $N_{pix}$ is the number of pixels.
This is the lowest possible complexity for a convolution operation without approximations (e.g. by sketching).
The flexibility of modeling the data domain with a graph allows to easily model data that spans only a part of the sphere, or data that is not uniformly sampled.
Our hierarchical pooling strategy exploits a hierarchical pixelisation of the manifold to analyse the data at multiple scales.
% Such hierarchical pixelisations of the sphere have been well studied.

@@

As there is no notion of translation on a graph, we cannot convolve two graph signals in a strict sense.
We can, however, convolve a signal with a kernel defined in the spectral domain.
More precisely, we can filter a graph signal by a kernel.
Given the convolution kernel
$h: \R_+ \rightarrow \R$, a signal $\f \in \R^{N_{pix}}$ on the graph is filtered as
\begin{equation} \label{eqn:graph_convolution}
	h(\L) \f = \U h(\bLambda) \U\trans \f,
\end{equation}
where $h(\bLambda)$ is a diagonal matrix where $h(\bLambda)_{ii} = h(\bLambda_{ii})$, and $h(\bLambda) \hat{\f}$ is the spectral representation of $\f$ after filtering.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{filtering_speed}
    \caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes.
    The fast spherical harmonic transform (SHT) is implemented by the \texttt{healpy} Python package (via the \texttt{healpy.sphtfunc.smoothing} function with $\ell_{max} = 3 N_{side} - 1$, the default).
    The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the \texttt{numpy} and \texttt{scipy} Python packages.
    Both are executed on a single core.
    The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3)$ for the spherical harmonics, where $\ell_{max}$ is the largest degree (i.e., the spectrum is truncated at $\ell_{max}$).}
    \label{fig:filtering_speed}
\end{figure}

While \eqnref{graph_convolution} is a theoretically well justified definition of the convolution of a graph signal $\f$ by a kernel $h$, it is computationally cumbersome.
As no general Fast Fourier Transform (FFT) exist for graphs, the execution of the Fourier transform by multiplication of the signal $\f$ by the dense matrix $\U$ costs $\bO(N_{pix}^2)$ operations.
% \footnote{See \citep{le2018fgft} for a comparison between the polynomial definition of filters presented here and a class of fast graph Fourier transforms based on the approximation of the Fourier basis $\U$ by a product of sparse matrices.}
%\TK{I would just cite the paper, we have not introduced the polynomials at this stage yet.}
In a CNN, this operation has to be performed for each training step: forward and backward.
% As current training procedures are not data-efficient and require a lot of samples, a quadratic computational cost in the sample size is a .
%\TK{what does "data-efficient mean"? removing for now}
As current training procedures require processing of many samples, it would be very slow to perform them for a quadratic computational cost in the sample size.
Moreover, the eigen-decomposition of the Laplacian $\L$ is needed to obtain the Fourier basis $\U$ in the first place.
That has a unique cost of $\bO(N_{pix}^3)$ operations.

Fortunately, both of these computational issues can be overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
In that specific case, the filtering operation \eqnref{graph_convolution} becomes
\begin{equation} \label{eqn:graph_convolution_polynomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
an operation requiring $k$ multiplications of the graph Laplacian matrix $\L$.
In our case, the Laplacian is extremly sparse, with $8-9$ entries per line, making the filtering operation for a fixed $K$ linear in the number of pixel $N_{pix}$.
% Moreover, here happens the transition from filtering in the spectral domain to filtering in the vertex domain.
This operation is performed in the vertex domain.
The matrix $\L^k$ captures $k$-neighborhoods of each pixel, i.e., $(\L^k)_{ij}$ is non-zero if and only if nodes $v_i$ and $v_j$ are connected by a path of length $k$.
As such, filtering with a polynomial convolution kernel can be interpreted in the pixel (vertex) domain as a weighted linear combination of the pixels neighbors values. The weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean signal-domain convolution to a patch. In the classical setting, the convolution with a patch is simply a weighted sum of the neighbors with the weights defined by the patch. Similarly, each lines of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an "irregular patch of radius $K$" for our irregular domain.
% \todo{Better vertex domain explanation. 
% \nati{I gave it a try. Should we take a single vertex and explain what heppens to it? Note that I suppressed the following paragraph. }}
% \TK{For me it's understandable:)}

@@

Following \citet{defferrard2016convolutional}, we define our filters as Chebyshev polynomials such as the convolution operation becomes
\begin{equation} \label{eqn:graph_convolution_cheby}
	h_\theta\left(\tL\right) \f = \U \left(\sum_{k=0}^K \theta_k T_k\left(\tilde{\bLambda}\right) \right) \U\trans \f = \sum_{k=0}^K \theta_k T_k\left(\tL\right) \f,
\end{equation}
where $\tL = \frac{2}{\lambda_{max}} \L - \b{I} = -\frac{2}{\lambda_{max}} \D^{-1/2} \W \D^{-1/2}$ is the rescaled Laplacian with eigenvalues $\tilde{\b \Lambda}$ in $[-1, 1]$ and $T_k(\cdot)$ is the Chebyshev polynomial of degree $k$, defined by the recursive relation $T_k(\tL) = 2\tL T_{k-1}(\tL) - T_{k-2}(\tL)$, $T_1(\tL) = \tL$, $T_0(\tL) = \b{I}$. While definitions \eqnref{graph_convolution_polynomial} and \eqnref{graph_convolution_cheby} both allow the representation of the same set of filters, Chebyshev polynomials are preferred as they form an orthogonal basis on the interval $[-1, 1]$. Moreover, they factor out $\tL^{k-2}$ from $T_k(\tL)$, e.g., the 3-neighborhood is not mixed with the 5-neighborhood. That results in easier learning when optimizing the parameters $\theta$.
Finally, the coefficients can be computed such as the truncated Chebyshev expansions gives an approximate minimax polynomial of a given filter \citep{hammond2011wavelets}. While that is irrelevant when learning filters, it is useful when smoothing a signal on the sphere by filtering it with an approximated low-pass filter.
In practice, we found out in our experiment that Chebyshev polynomials are slightly more stable than the monomials used in~\cite{khasanova2017graphomni}(i.e. optimizing directly the $\theta_k$ in \eqnref{graph_convolution_polynomial}).
We believe that this is due to their orthogonality (as explained above) and uniformity\footnote{The amplitude of Chebyshev polynomials is more or less constant over the domain $[-1, 1]$ independently of the order $k$. On the contrary the amplitude of $x^k$ is very different for $|x|\approx 0$ and $|x| \approx 1$.}.
% \nati{I added a sentence and I think we should not give much more details here. If we do an experiment, we can comment in the experiment section.}
% \todo{(i) scale of $\theta_k$: they are mostly uniform for Chebyshev polynomials, end exploding for monomials. Though SGD implicitly imposes a uniform prior on the learned parameters.}
% \todo{(ii) numerical stability of $\L^k \f$ v.s. $T_k(\L) \f$}
% \todo{(iii) in the code, initialize $\theta_k$ with stdev adapted to $F_{in}$ and $K$}

Exploiting the recursive formulation of Chebyshev polynomials, evaluating \eqnref{graph_convolution_cheby} requires $\bO(K)$ multiplications of the dense vector $\f$ with the sparse matrix $\tL$.
The cost of one such multiplication is $\bO(|\E|)$. By construction of our graph, $|\E| < 8 N_{pix}$ and the overall computational cost of the convolution is thus reduced to $\bO(N_{pix})$ operations.
That is much more efficient than filtering with spherical harmonics, even tho the HEALPix was designed as an iso-latitude sampling to leverage a fast spherical transform.
That is especially true for smooth kernels which require a low polynomial degree $K$.
\figref{filtering_speed} demonstrates the speed comparison of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here.
The performance of our approach is ten to twenty times faster for $N_{side} = 2048$, with $K=20$ and $K=5$, respectively.

@@

\citet{boomsma2017spherical} propose to apply a standard 2D CNN to a grid discretization of the sphere. Two pixelisations, illustrated in \figref{sphere_grids}, were used: the equirectangular grid, and the cubed-sphere grid defined by~\citet{ronchi1996cubed}.
As this formulation uses the standard 2D convolutions, all the optimizations developed for 2D images can be applied, what make it computationally efficient.
Some losses are due to the handling of boundary conditions, for example by padding a grid with the content of the opposite side (equirectangular) or of the neighboring grids (cubed-sphere, HEALPix).
%It is, however, restricted to those grid sampling sets.
As many pixelizations are based on the subdivision of a base polyhedron
Similarly to the cubed-sphere, HEALPix is made of twelve grids. As those grids are not independent, each grid has to be  to handle , .
% Furthermore, the grid requirement makes it impossible for the convolution to capture the spherical properties of the domain, i.e., the cubed-sphere sampling is by definition adapted to a cube and not to a sphere. Nope, the graph can adapt to any sampling.
% Moreover, those samplings are not ideal as their resolution is not constant across the sphere.

% Moreover, the pixel area is not constant across the sphere in those sampling schemes, which may limit their potential applications.
% Indeed, the singularities, such as the poles for the equirectangular grid, and the corners of the cube for the cubed-sphere, are over-sampled.
% Equal areas of discrete elements is advantageous for many physical measurements, as white noise at the sampling frequency of the instrument gets integrated exactly into white noise in the pixel space, and sky signals are sampled without regional dependence \citep{gorski2005healpix}.
% => 2D CNN can be applied on HEALPix as well.

% \nati{Michael: Under some very specific hypothesis, this second approach is a particular case of our method, i.e. (i) our method with a stupid sampling, (ii) assumption that graph convolution on a grid == 2d convolution. Do you think we should mention that? I think we should not.}
Finally, a 2D CNN learns directional filters, which are well defined with respect of the image coordinates.
Directions that make much sense for images (up, down, left, and right) are arbitrarily defined when setting the origin of the grids on the sphere.
Any convolution using directions on the sphere is not equivariant to rotation.

% As presented by~\cite{cohen2018spherical}, the most mathematical approach to extend the convolution on a sphere is to use a spherical Fourier transform. The convolution is then simply defined as the product in the spectral domain. This approach requires one Fourier and one inverse Fourier transform per convolution which remain, even with accelerated algorithms, expensive.

>>>>

\section*{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
	\item (TK) Related work in cosmo. The only cosmological applications of spherical CNNs in the related work are about projection on the tangent plane \citep{schmelze2017cosmologicalmodel, fluri2018deep, gupta2018nongaussianinformation}. Aren't there applications in cosmology for the other (c.f. related work section) approaches? Most of them are illustrated in omnidirectional imagery.
	\item (MD) distinction between CNN and FCN
    \item (done!) (NP) make figure A.13 remove white space
    \item (first version done!)(NP+MD) paragraph on our chosen architecture in the Experiments sections
    \item (Done) (NP) vertex interpretation
    \item (MD) Are you happy with the vertex interpretation in section 2.5 Efficient convolutions? Also do you like the end of 2.4?
    \item (NP) Finalise readme on Github code
    \item branch with notebooks with plots (results) NP
    \item NP move the repo to HealpixNet
    \item check if papers which appear as arxiv were not published in a conference or journal
    \item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings. \nati{I think this is too much for this paper.}
    \item TK check part sphere notebook and whole sphere
    \item (done!) (TK) update the README for data zip - file from NP
    \item (almost done) TK read everything
    \item (done!) (TK) uniformize text size in figures
    \item (done!) (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook. \nati{We need to talk about that.}
    \item (done!) (TK) re-make figures to make the labels/legend/titles the same font size as the main text
    \item (done!) TK check the appendix
    \item (done!) TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item (done!) (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (done!) TK read the result section
    \item (done!) (TK) add reference to figure (pooling) to the area of orders
    \item (done!) (TK) add section in conclusions/ paragraph on possible extensions to the code: dense output, regression, GANs, segmentation
	\item Beware to not confuse reader with synonyms, e.g. pixelization and sampling.
    \item Check names: HealpixNet, HealpixCNN, HealpixFCN
    \item Be sure that the convolution on SO(3) is clear: we want to say (a) is the one associated with the SO(3) rotation group (b) can be performed  in the spectral domain by a multiplication (c) is equivariant to rotation (d) is described well in Section 2 of \citep{kondor2018clebsch}.
	\item Have a "Contributions" paragraph?
	\item Notation: write element $i$ of vector $y$ as $y_i$, $y(i)$, or $y[i]$? Be consistent. Same for matrices.
    % \item (TK took exec decision and moved it :D) Should the heat example go to appendix?
    % \item (done! first version done) conclusions
    % \item (done!) Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    % \item (done!) (TK NP) if we have time make a new candidate for network plot (optional)
    % \item (done!) TK read the heat section again
    % \item (done!) (TK) make a new proposition for a plot Figure 6: Two levels of coarsening and pooling
    % \item \todo{Notation GC32 - P256 - FC10 to describe an architecture? Maybe more useful in Experiments where we describe the architecture used for each experiment.}
    % Nati: I did it without this notation. Since we have two architectures, I think it is OK.
\end{itemize}

@@

% \todo{@Tomek: can you explain why we care about parts of the sphere? (that is an advantage of the graph model over the spherical harmonics)}
% \todo{@Tomek: How is the data acquired? Do telescopes and satellites take measurements according to the HEALPix sampling? (it is also an advantage of our method to be agnostic to the sampling, which might be useful to researchers working with missing data or different models)}
% \nati{@Tomek: I think what Michael has in mind is that the graph approach is very versatile and extendable. It can be used with other samplings (even irregular ones) or only part of the sphere. This flexibility might be an advantage to mention. In particular, compared to~\citet{cohen2018sphericalcnn} it allows us to do only do the required computation when only a part of the sphere is considered.}

So far these algorithms have mostly been demonstrated on Euclidean domains, such as flat images.
Perhaps the most straightforward approach to the analysis of spherical data is to divide it into small chunks and project those on flat 2D surfaces. \todo{Tune according to the story in the related work.}
While this approach has been demonstrated to work \citep{fluri2018deep,gupta2018nongaussianinformation,schmelze2017cosmologicalmodel,gillet2018deeplearning}, it may be more natural for some applications and accurate to work directly on the sphere.
The main challenge in designing a CNN on the sphere is to define a convolution operation that is suitable for this irregularly sampled domain.
To be able to train the network efficiently, the convolution has to be performed fast.
Another important feature is for a CNN to be able to work on part-spheres, as for many cosmological observations cover only a part of the sky.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location, and for space-based instruments due to masking of the galactic plane area.

% cohen2017convolutional (ICML workshop) and cohen2018sphericalcnn (ICLR best paper) are redundant
A first attempt to generalize CNNs to spherical data was to apply a standard 2D CNN to a grid discretisation of the sphere~\citep{boomsma2017spherical}. While this approach uses the well-developed 2D convolution and hierarchical pooling, it is restricted to those specific grid samplings which are highly irregular on the sphere.
An attempt by~\citet{cohen2018sphericalcnn} was to leverage the spherical Fourier transform and to perform the SO(3) convolution in the spectral domain, thanks to the convolution theorem.
The resulting convolution is rotational equivariant, meaning that a rotation of the input implies the same rotation of the output.
While there exists fast spherical Fourier transforms for some samplings, this approach can still be computationally expensive.
Finally, the use of a graph to model the discretised sphere was considered by~\citet{khasanova2017graphomni} as well, towards the goal of designing network that were invariant to rotations.
While their architecture accommodates these invariances on the sphere, they did not take advantage of a hierarchical pixelisation.
Moreover, the statistical layer, which introduces the invariance, needs to be hand-crafted to capture statistics of interest.
\todo{better identify the shortcomings: (i) part of sphere / missing data, (ii) computational efficiency, (iii) equivarience}

To overcome those shortcomings, we introduce an efficient spherical CNN formulation which uses (i) convolutions on graphs and (ii) hierarchical pooling.
The main idea is to represent the sphere $S^2$, a 2-dimensional manifold embedded in $\mathbb{R}^3$, as a graph of connected pixels.
The length (or weight) of the shortest path between two pixels is an approximation of the geodesic distance between them on the manifold.
The use of this approach is enabled by recent developments in graph signal processing, where fundamental tools, such as for example the Fourier transform, were defined for graphs \citep{shuman2013emerging}.
Our approach uses the graph CNN formulation introduced by \citet{defferrard2016convolutional}, which can accommodate any sampling of the sphere.
Using a $k$-nearest neighbours graph, the convolution operation costs $\mathcal{O}(N_{pix})$ operations, where $N_{pix}$ is the number of pixels.
This is the lowest possible complexity for a convolution operation without approximations (e.g., by sketching).
The flexibility of modeling the data domain with a graph allows to easily model data that spans only a part of the sphere, or data that is not uniformly sampled.
Our hierarchical pooling strategy exploits a hierarchical pixelisation of the manifold to analyse the data at multiple scales.
% Such hierarchical pixelisations of the sphere have been well studied.

@@

As there is no notion of translation on a graph, we cannot convolve two graph signals in a strict sense.
We can, however, convolve a signal with a kernel defined in the spectral domain.
More precisely, we can filter a graph signal by a kernel.
Given the convolution kernel
$h: \R_+ \rightarrow \R$, a signal $\f \in \R^{N_{pix}}$ on the graph is filtered as
\begin{equation} \label{eqn:graph_convolution_fourier}
	h(\L) \f = \U h(\bLambda) \U\trans \f,
\end{equation}
where $h(\bLambda)$ is a diagonal matrix where $h(\bLambda)_{ii} = h(\bLambda_{ii})$, and $h(\bLambda) \hat{\f}$ is the spectral representation of $\f$ after filtering.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{filtering_speed}
    \caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes.
    The fast spherical harmonic transform (SHT) is implemented by the \texttt{healpy} Python package (via the \texttt{healpy.sphtfunc.smoothing} function with $\ell_{max} = 3 N_{side} - 1$, the default).
    The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the \texttt{numpy} and \texttt{scipy} Python packages.
    Both are executed on a single core.
	The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3) = \bO(N_{pix}^{3/2})$ for the spherical harmonics, where $\ell_{max}$ is the largest angular frequency.}
    \label{fig:filtering_speed}
\end{figure}

While \eqnref{graph_convolution_fourier} is a theoretically well justified definition of the convolution of a graph signal $\f$ by a kernel $h$, it is computationally cumbersome.
As no general Fast Fourier Transform (FFT) exist for graphs, the execution of the Fourier transform by multiplication of the signal $\f$ by the dense matrix $\U$ costs $\bO(N_{pix}^2)$ operations.
% \footnote{See \citep{le2018fgft} for a comparison between the polynomial definition of filters presented here and a class of fast graph Fourier transforms based on the approximation of the Fourier basis $\U$ by a product of sparse matrices.}
%\TK{I would just cite the paper, we have not introduced the polynomials at this stage yet.}
In a CNN, this operation has to be performed for each training step: forward and backward.
% As current training procedures are not data-efficient and require a lot of samples, a quadratic computational cost in the sample size is a .
%\TK{what does "data-efficient mean"? removing for now}
As current training procedures require processing of many samples, it would be very slow to perform them for a quadratic computational cost in the sample size.
Moreover, the eigen-decomposition of the Laplacian $\L$ is needed to obtain the Fourier basis $\U$ in the first place.
That has a unique cost of $\bO(N_{pix}^3)$ operations.

Fortunately, both of these computational issues can be overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
In that specific case, the filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
an operation requiring $k$ multiplications of the graph Laplacian matrix $\L$.
In our case, the Laplacian is extremly sparse, with $8-9$ entries per line, making the filtering operation for a fixed $K$ linear in the number of pixel $N_{pix}$.
% Moreover, here happens the transition from filtering in the spectral domain to filtering in the vertex domain.
This operation is performed in the vertex domain.
The matrix $\L^k$ captures $k$-neighborhoods of each pixel, i.e., $(\L^k)_{ij}$ is non-zero if and only if nodes $v_i$ and $v_j$ are connected by a path of length $k$.
As such, filtering with a polynomial convolution kernel can be interpreted in the pixel (vertex) domain as a weighted linear combination of the pixels neighbors values. The weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean signal-domain convolution to a patch. In the classical setting, the convolution with a patch is simply a weighted sum of the neighbors with the weights defined by the patch. Similarly, each lines of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an "irregular patch of radius $K$" for our irregular domain.
% \todo{Better vertex domain explanation.
% \nati{I gave it a try. Should we take a single vertex and explain what heppens to it? Note that I suppressed the following paragraph. }}
% \TK{For me it's understandable:)}

@@

Following \citet{defferrard2016convolutional}, we define our filters as Chebyshev polynomials such as the convolution operation becomes
\begin{equation} \label{eqn:graph_convolution_cheby}
	h_\theta\left(\tL\right) \f = \U \left(\sum_{k=0}^K \theta_k T_k\left(\tilde{\bLambda}\right) \right) \U\trans \f = \sum_{k=0}^K \theta_k T_k\left(\tL\right) \f,
\end{equation}
where $\tL = \frac{2}{\lambda_{max}} \L - \b{I} = -\frac{2}{\lambda_{max}} \D^{-1/2} \W \D^{-1/2}$ is the rescaled Laplacian with eigenvalues $\tilde{\b \Lambda}$ in $[-1, 1]$ and $T_k(\cdot)$ is the Chebyshev polynomial of degree $k$, defined by the recursive relation $T_k(\tL) = 2\tL T_{k-1}(\tL) - T_{k-2}(\tL)$, $T_1(\tL) = \tL$, $T_0(\tL) = \b{I}$. While definitions \eqnref{graph_convolution_monomial} and \eqnref{graph_convolution_cheby} both allow the representation of the same set of filters, Chebyshev polynomials are preferred as they form an orthogonal basis on the interval $[-1, 1]$. Moreover, they factor out $\tL^{k-2}$ from $T_k(\tL)$, e.g., the 3-neighborhood is not mixed with the 5-neighborhood. That results in easier learning when optimizing the parameters $\theta$.
Finally, the coefficients can be computed such as the truncated Chebyshev expansions gives an approximate minimax polynomial of a given filter \citep{hammond2011wavelets}. While that is irrelevant when learning filters, it is useful when smoothing a signal on the sphere by filtering it with an approximated low-pass filter.
In practice, we found out in our experiment that Chebyshev polynomials are slightly more stable than the monomials used in~\cite{khasanova2017graphomni}(i.e. optimizing directly the $\theta_k$ in \eqnref{graph_convolution_monomial}).
We believe that this is due to their orthogonality (as explained above) and uniformity\footnote{The amplitude of Chebyshev polynomials is more or less constant over the domain $[-1, 1]$ independently of the order $k$. On the contrary the amplitude of $x^k$ is very different for $|x|\approx 0$ and $|x| \approx 1$.}.
% \nati{I added a sentence and I think we should not give much more details here. If we do an experiment, we can comment in the experiment section.}
% \todo{(i) scale of $\theta_k$: they are mostly uniform for Chebyshev polynomials, end exploding for monomials. Though SGD implicitly imposes a uniform prior on the learned parameters.}
% \todo{(ii) numerical stability of $\L^k \f$ v.s. $T_k(\L) \f$}
% \todo{(iii) in the code, initialize $\theta_k$ with stdev adapted to $F_{in}$ and $K$}

Exploiting the recursive formulation of Chebyshev polynomials, evaluating \eqnref{graph_convolution_cheby} requires $\bO(K)$ multiplications of the dense vector $\f$ with the sparse matrix $\tL$.
The cost of one such multiplication is $\bO(|\E|)$. By construction of our graph, $|\E| < 8 N_{pix}$ and the overall computational cost of the convolution is thus reduced to $\bO(N_{pix})$ operations.
That is much more efficient than filtering with spherical harmonics, even tho the HEALPix was designed as an iso-latitude sampling to leverage a fast spherical transform.
That is especially true for smooth kernels which require a low polynomial degree $K$.
\figref{filtering_speed} compares the speed of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here.
A naive implementation of our method is ten to twenty times faster for $N_{side} = 2048$, with $K=20$ and $K=5$, respectively, than using the spherical harmonic transform implemented by the highly optimized library used by HEALPix, \pkg{libsharp} \citep{reinecke2013libsharp}.

@@

A first approach, explored by \citet{boomsma2017spherical} for molecular modeling and \citet{su2017sphericalconv, coors2018spherenet} for omnidirectional imaging, is to use a 2D CNN on a discretization of the sphere that is a grid, such as the equirectangular projection (\figref{sphere_grids} left), or a set of grids, such as the cubed-sphere defined by~\citet{ronchi1996cubed} (\figref{sphere_grids} middle).
As this formulation uses the standard 2D convolution, all the optimizations developed for images can be applied, what make it computationally efficient.
This approach is applicable to the many pixelizations that are based on a regular subdivision of a base polyhedron, such as HEALPix. Each base polyhedron then forms a grid.
Care has to be taken to handle boundary conditions: for example by padding a grid with the content of the opposite side (equirectangular) or of the neighboring grids (cubed-sphere, HEALPix). That incures some computational losses.
For samplings that are not equal area, such as the equirectangular projection, the convolution operation can be adjusted to take the induced distortion into account \citep{su2017sphericalconv, coors2018spherenet}.
% Furthermore, the grid requirement makes it impossible for the convolution to capture the spherical properties of the domain, i.e., the cubed-sphere sampling is by definition adapted to a cube and not to a sphere. Nope, the graph can adapt to any sampling.

Another approach to leverage 2D CNNs is to project spherical data onto many tangent planes, which are flat 2D surfaces. This approach has been extensively used for cosmological maps \citep{fluri2018deep, gupta2018nongaussianinformation, schmelze2017cosmologicalmodel, gillet2018deeplearning} and omnidirectional imaging \citep{xiao2012recognizing, zhang2014panocontext}. This idea has been generalized to arbitrary 2D manifolds for shape alignment and retrieval \citep{masci2015gcnn, boscaini2016acnn, monti2017monet}.

The main issue with the above two approaches is that they depend on a (local) coordinate system to define anisotropic filters, i.e., filters which are direction dependent. Direction is well defined and matters for some applications, such as the analysis of weather and climate data on the Earth (north, south, east, west), and omnidirectional imaging (up, down). Indeed, full rotation invariance has been shown to reduce discriminative power for omnidirectional imaging \citep{coors2018spherenet}. Some problems are however intrinsically invariant to rotation, such as the analysis of cosmological maps, or the modeling of atoms and molecules. In such cases, directions are arbitrarily defined when setting the origin of the pixelization. A rotation equivariant convolution, i.e., a rotation of the input implies the same rotation in the output, thus has to be isotropic.

% Moreover, when different grids require different filters as their orientation is not compatible.
% => same as for tangent planes
